Explore the SingleR results

### 2025-07-14 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/singler-results.Rmd
Very WIP: expression plot prep

>>>>

```{r}
# this time we are not considering putative tumor
direct_celltype_matches <- c(
  "B cell",
  "T cell",
  # note that we need to collapse the consensus to compare
  "myeloid",
  "fibroblast",
  "endothelial cell",
  "plasma cell",
  "natural killer cell", 
  "unknown_consensus" # we do want to keep these for this analysis
)

situation_df <- celltype_recoded_df |>
  # recode labels for comparison
  dplyr::mutate(
    label_recoded = ifelse(
      label_recoded %in% consensus_myeloid,
      "myeloid",
      label_recoded
  )) |>
  dplyr::filter(label_recoded %in% direct_celltype_matches) |>
  dplyr::select(cell_id, annotation_type, label_recoded) |>
  # make it wider to label the situation
  tidyr::pivot_wider(
    names_from = "annotation_type",
    values_from = "label_recoded"
  ) |>
  # remove NAs:
  # - consensus NAs are those with other cell types not to include here
  # - singler NAs were unclassified cells
  tidyr::drop_na() |>
  dplyr::mutate(
    annotation_situation = dplyr::case_when(
      consensus_validation == "unknown_consensus" ~ "consensus not classified", 
      consensus_validation == singler ~ "matching annotations",
      consensus_validation != singler ~ "different annotations"
    )
  ) |>
  # make it long again
  tidyr::pivot_longer(
    c(consensus_validation, singler),
    names_to = "annotation_type", 
    values_to = "label_recoded"
  )

cells <- unique(situation_df$cell_id)
genes <- unique(consensus_markers_df$ensembl_gene_id)
```

```{r}
scuttle::makePerCellDF(
    merged_sce[genes, cells],
    features = genes,
    assay.type = "logcounts",
    use.coldata = "cell_id",
    use.dimred = FALSE
  ) |>
    tidyr::pivot_longer(starts_with("ENSG"), names_to = "ensembl_gene_id", values_to = "logcounts") |>
  dplyr::left_join(situation_df, by = "cell_id") |>
  dplyr::group_by(label_recoded, marker_gene_label, ensembl_gene_id) |>
  dplyr::summarize(
    detected_count = sum(detected),
    mean_exp = mean(logcounts)
  ) |>
  dplyr::ungroup()

ggplot(situation_df) + 
  aes(x = singler, y = )

  group_stats_df <- all_info_df |>
    # remove genes that aren't present in final annotations
    dplyr::filter(gene_symbol %in% marker_gene_order) |>
    # for each assigned cell type/marker gene combo get total detected and mean expression
    dplyr::group_by(label_recoded, marker_gene_label, ensembl_gene_id) |>
    dplyr::summarize(
      detected_count = sum(detected),
      mean_exp = mean(logcounts)
    ) |>

```
